### 2009-11-23 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revisions for chapters 2 and 3 after prof. Puder's reviews.
- Removing blurried images;
- Adding the data provenance taxonomy

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@656 559edde2-a4d4-44e0-96fc-4035e40d7396

This chapter will first give an overview on different related problems
described  in the literature that solved problems related to data modeling and 
representation in sensor networks. Then, it presents the design, architecture 
and implementation details of the DSP Platform, which is the bases for this 
dissertation. Finally, this chapter will briefly give a tutorial on advanced 
topics such as the differences between the KVP data model and the relational 
data model, as well as discussing about the approaches used on Database
Sharding.

@@

Sensor Networks are specialized types of network systems comprised of network 
nodes with specific goals to observe and collect data. The following
subsections describe classes of characteristis of these types of sensor
networks, with applications on research, industrial or militery.

@@

\section{How to describe collected sensor data using Data Provenance}
\label{sec:sn-provenance}

>>>>

This chapter will first give an overview on different related problems
described in the literature that solved problems related to data modeling and
representation in sensor networks.

@@

Sensor Networks are specialized types of network systems comprised of network 
nodes with specific goals to observe and collect data. Data Persistence in
Sensor Networks, however, depends on the approaches used to store the data,
since data nodes can be located close to the devices in the network or in a
centralized node. In this way, the query process can take place on the network
nodes themselves or on the centralized data sink. The former approach is
generally used when the data is collected at the sensor node on its local
storage \cite{sn-storage01}\cite{sn-storage03}, strategy mainly used to
maximize the sensor's resources while minimizing energy consumption
\cite{sn-storage04}. However, this strategy is used when data archival is not
the primary reason, given the limited resources on the sensor devices. For
this reason, the the latter strategy has been used when the collected data is
primarily archived in a centralized database system for later reuse
\cite{sn-storage02}.

When it comes to data representation, the relational data model
\cite{relational-model} is one of the most used approaches to represent data
in sensor network databases in any of the query processing strategies and, in
most cases, with a modified version of the SQL language \cite{sn-db-newop}.
However, the design of the database system must be done with prior knowledge
of the sensor types when choosing the relational data model, as the database
design must be normalized \cite{db-normalization} for the entities already
identified for the system. For instance, the introduction of a new sensor
device might represent a potential change on the database structure by seeking
a new normalized version of the database structure. In addition to challenges
of maintaining the database schema normalized, the use of Data Provenance
approaches in sensor network data \cite{sn-provenance} usually addresses the
problem related with the lineage of data  and its full history. However, once
the data model is chosen and implemented, one of the most important
non-functional requirements in sensor networks is scalability of the
persistence storage. Since an operational sensor network can produce huge
amounts of data, it must be able to cope with the increase and data load
during data gathering, and yet provide users the same performance during data
retrieval by using different strategies such a data-centric storage
\cite{sn-storage03}.

The following subsections describe classes of characteristis of these types of
sensor networks, with applications on research, industrial or militery.

@@

\section{Data Provenance: How to describe collected data?}
\label{sec:sn-provenance}

### 2009-11-23 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revisions for chapters 2 and 3 after prof. Puder's reviews.
- Removing blurried images;
- Adding the data provenance taxonomy

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@656 559edde2-a4d4-44e0-96fc-4035e40d7396

Different mechanisms are used for persisting data collected from sensor
networks, as described by the section regarding the state of the art in sensor
networks and different approaches for persistence. This section analyzes these
approaches and presents the different approaches through taxonomy
\footnote{Taxonomy is the practice and science of classification, by using
taxonomic units known as taxa (singular taxon)}, which will be used throughout
this work to evaluate different existing database systems implementations,
taking into account different requirements and specifications of different
sensor networks.

@@

@@

No matter the type of data model used, the support for Data Provenance as shown
in section \ref{sec:sn-provenance} is as important as the data being
collected. For this reason, the data model taxonomy must have support for the
data types that support the use of provenance-related attributes in the
collected sensor data.

>>>>

Different mechanisms are used for persisting data collected from sensor
networks, as described by the section regarding the state of the art in sensor
networks and different approaches for persistence. This section analyzes these
approaches and presents the different approaches through 
taxonomy\footnote{Taxonomy is the practice and science of classification, by
using taxonomic units known as taxa (singular taxon)}, which will be used
throughout this work to evaluate different existing database systems
implementations, taking into account different requirements and specifications
of different sensor networks.

@@

No matter the type of resource management used, the first taxonomy presented is
related to the purpose of the use of the collected sensor data. That is, when
and how the collected data is going to be used as described in section
\ref{sec:sn-data-purpose}. In addition to it, the mechanism for data storage
described in section \ref{sec:sn-infrastructure} can be related to each of
them. The related taxon is depicted by figure \ref{fig:taxonomy-data-location}.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-data-location}
  \caption{The Nature and Location of Sensor Data Taxonomy}
  \label{fig:taxonomy-data-location}
\end{figure}

@@

\section{Data Provenance Taxonomy}

No matter the type of data model used, the support for Data Provenance as shown
in section \ref{sec:sn-provenance} is as important as the data being collected.
For this reason, the data model taxonomy must have support for the data types
that support the use of provenance-related attributes in the collected sensor
data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the data types
for the metadata support.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-data-provenance}
  \caption{Data Provenance Taxonomy}
  \label{fig:taxonomy-data-provenance}
\end{figure}

\begin{itemize}
  \item \textbf{What: Data Identity}: the data type that uniquely identifies
  the collected sensor dat;
  \item \textbf{When: Time Dimension}: the data type that identifies the
  collected data;
  \item \textbf{Where: Data Location}: an optional data type that identifies
  where the data came from.
\end{itemize}
